---
title: 'Bagged Topic models: Simulation 1'
author: "Kushal K Dey"
date: "March 6, 2016"
output: html_document
---

## Overview

We want a faster way of doing classification using topic models. Usually there are lots of genes/ features in the dataset but only a handful of them drive the classification. Our aim would be to choose variables judiciously and also fit the `classtpx` models for small batches of variables at a time, build an ensemble and then average over the results to get the final output.

## Simulation Set up

```{r echo=TRUE, eval=TRUE}
n.out <- 1000
omega_sim <- rbind( cbind( rep(1, n.out), rep(0, n.out)), 
                    cbind( rep(0, n.out), rep(1, n.out)),
                    cbind( seq(0.6, 0.4, length.out = n.out), 
                           1- seq(0.6, 0.4,length.out=n.out)) )
dim(omega_sim)

K <- dim(omega_sim)[2]
barplot(t(omega_sim), 
        col = 2:(K+1),
        axisnames = F, space = 0, 
        border = NA, 
        main=paste("No. of clusters=", K),
        las=1, ylim = c(0,1), cex.axis=1.5,cex.main=1.4)

freq <- rbind( c(0.1, 0.2, rep(0.70/998, 998)),
               c(rep(0.70/998,998), 0.1, 0.2) )
str(freq)

counts <- t( do.call(cbind,
                     lapply(1:dim(omega_sim)[1], 
                            function(x) 
                                rmultinom(1,1000,prob=omega_sim[x,]%*%freq))))
dim(counts)
```

We apply `classtpx` model on the counts data with the first 100 genes in one cluster and the genes from 1000 to 1100 in another cluster. 

```{r echo=TRUE, eval=TRUE}
library(classtpx)
known_indices <- c(1:100,1001:1100);
omega_known <- omega_sim[known_indices,]
system.time(Topic_clus <- classtpx::class_topics(
    counts, 
    K, 
    known_indices = known_indices, 
    omega_known = omega_known, 
    tol=0.001))
```

We plot the Structure plot of the omega matrix.

```{r echo=TRUE, eval=TRUE}
# These are estimated gene relative expression
docweights <- Topic_clus$omega

library(permute)
library(BioPhysConnectoR)

# Decide on the correspondance of the simulated
# and of the estimated relative expression matrix
# 
# We consider all possible cluster ordering in the 
# estimated relative expression matrix and select
# the relative expression matrix with the nearest
# distance to the simulate relative experssion matrix
# (forbieus norm)

perm_set <- rbind(1:K,allPerms(1:K))
diff <- array(0,dim(perm_set)[1]) 
for (p in 1:dim(perm_set)[1])
{
    temp <- docweights[, perm_set[p,]]
    diff[p] <- fnorm(temp,omega_sim)
}

p_star <- which(diff==min(diff))
docweights <- docweights[,perm_set[p_star,]]

barplot(t(docweights),
        col = 2:(K+1),
        axisnames = F, space = 0, border = NA, 
        main=paste("No. of clusters=", K),
        las=1, ylim=c(0,1), cex.axis=1.5, cex.main=1.4)
```

## Bagged topic models

We form $20$ batches of $50$ genes each and then for each batch, we fit the topic model in parallel.

```{r echo=TRUE, eval=TRUE}

num <- 50
system.time(bag_topics <- parallel::mclapply(1:20, function(l)
                           {
                                counts_temp <- counts[,((l-1)*num+1):(l*num)];
                                index <- which(rowSums(counts_temp)==0);
                                if(length(index)!=0){
                                  counts_temp[index,]=1;
                                }
                                out <- classtpx::class_topics(counts_temp,                                                                   K=2,
                                                 known_indices=known_indices,                                                  omega_known = omega_known,                                                    tol=0.1);
                                return(out)
                           }))

```

We build strength measures for the clusters. 

```{r echo=TRUE, eval=TRUE}
#barplot(t(bag_topics[[10]]$omega),
#        col = 2:(K+1),
#        axisnames = F, space = 0, border = NA, 
#        main=paste("No. of clusters=", K),
#        las=1, ylim=c(0,1), cex.axis=1.5, cex.main=1.4)

#CountClust::compare_omega(bag_topics[[1]]$omega[-known_indices,],omega_sim[-known_indices,])$cor.information_content

rep.row<-function(x,n){
   matrix(rep(x,each=n),nrow=n)
}
rep.col<-function(x,n){
   matrix(rep(x,each=n), ncol=n, byrow=TRUE)
}


omega_null <- rep.row(rep(1/K,K),dim(counts)[1])

#null_kl_rows <- rowSums(CountClust::compare_omega(bag_topics[[1]]$omega[-known_indices,],omega_null[-known_indices,])$kl.dist)
#bag_strength <- exp(-(max(null_kl_rows)/min(null_kl_rows)));

bag_strength_list <- sapply(1:20, function(l)
                                  {
                                             bag_strength_1 <- CountClust::compare_omega(bag_topics[[l]]$omega[-known_indices,],omega_sim[-known_indices,])$cor.information_content 
                                             suppressWarnings(null_kl_rows <- rowSums(CountClust::compare_omega(bag_topics[[l]]$omega[-known_indices,],omega_null[-known_indices,])$kl.dist))
                                              bag_strength_2 <- exp(-(max(null_kl_rows)/min(null_kl_rows)));
                                              bag_strength <- (1-bag_strength_2)*bag_strength_1;
                                              return(bag_strength)

                                  })

```

We just pool the batches with high values of cluster information and perform topic model and plot the omega proportion as Structure plot. 

```{r echo=TRUE, eval=TRUE}
counts_test <- counts[,c(1:50,950:1000)];
baggedTopic_clus <- classtpx::class_topics(counts_test,                                                                   K=2,
                                                 known_indices=known_indices,                                                  omega_known = omega_known,                                                    tol=0.1);

barplot(t(baggedTopic_clus$omega),
        col = 2:(K+1),
        axisnames = F, space = 0, border = NA, 
        main=paste("No. of clusters=", K),
        las=1, ylim=c(0,1), cex.axis=1.5, cex.main=1.4)
```

We further take a smaller subset of genes (assuming we had performed the gene level strength). Here we just assume the first 10 and the last 10 genes.

```{r echo=TRUE, eval=TRUE}
counts_test_2 <- counts[,c(1:10,990:1000)];
baggedTopic_clus_2 <- classtpx::class_topics(counts_test_2,                                                                   K=2,
                                                 known_indices=known_indices,                                                  omega_known = omega_known,                                                    tol=0.1);
barplot(t(baggedTopic_clus_2$omega),
        col = 2:(K+1),
        axisnames = F, space = 0, border = NA, 
        main=paste("No. of clusters=", K),
        las=1, ylim=c(0,1), cex.axis=1.5, cex.main=1.4)

```

